---
title: "Exercises"
author: "Nicolas Kolling Ribas"
date: "2023-01-25"
output: pdf_document
---

## 6.1 The CPU time and disk I/O’s of seven programs are shown in Table 6.7. Determine the equation for principal factors.

Load the dataset
```{r}
dataset <- read.csv("table-6.7.csv")
print(dataset)
```


Run principal-component analysis algorithm
```{r}
pca_result <- prcomp(dataset[3:4], scale = TRUE)
print(pca_result)
```

Plot the values of principal factors
```{r, message = FALSE}
library(ggfortify)
autoplot(pca_result, data = dataset, colour = 'Function', size = 3)
```

## 6.2 Using a spanning-tree algorithm for cluster analysis, prepare a dendrogram for the data shown in Table 6.7. Interpret the result of your analysis.

Make program name the row name, remove "Function" collumn
```{r}
library(tibble)
dataset <- column_to_rownames(dataset, var = "Program.Name")[,-1]
print(dataset)
```

Compute the intercluster distance matrix and do the agglomerative hierarchical clustering
```{r}
dist <- dist(scale(dataset), method = "euclidean")
clust <- hclust(dist, method = "ward.D2")
```

Plot the dendrogram
```{r}
dend <- as.dendrogram(clust)
plot(dend)
```
